---
title: "A1_Gillette_20204160.Rmd"
author: "EGillette (20204160)"
date: "2025-01-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

https://github.com/EmiliagG/BIOL432_Assignment1
This section involves loading in the measurement data
```{r}
  measurementData<-read.table("https://raw.githubusercontent.com/EmiliagG/BIOL432_Assignment1/main/measurements.csv", header=T, sep="," )
```

#Sorting the Data

The data is first sorted by species, then observer, and then limb volume

```{r}
arrange(measurementData, Organism, Observer, Volume, by.group=FALSE)
```
#Creating a Table for Average Limb Volume per Species

```{r}
measurementData %>%
  group_by(Organism) %>%
  summarize(Mean_volume= mean(Volume,na.rm = TRUE))
  
```
#Creating a Table for the Average Number of Observations




## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
